---
title: R Notebook
author: Yuan Tian
date: '2019-08-16'
slug: r-notebook
categories:
  - r
tags: []
image:
  caption: ''
  focal_point: ''
output: 
 blogdown::html_page:
  number_sections: true
  toc: true
  toc_depth: 3
  css: "/css/custom.css"
---

# **R Setup 101**

## R Packages for Basic R Installation or Update

For new installation, here is a list of most-used packages to install: 

```{r eval=FALSE}
install.packages('rmarkdown')
install.packages("tinytex")
# install TinyTeX (must-do-this-step)
tinytex::install_tinytex()  
install.packages('tidyverse')
install.packages('ggplot2')
#table creation with specific formatting
install.packages('kableExtra')
#for writing reproducible research paper in R Markdown
install.packages(c("stargazer","plotly","knitr","bookdown"))
#To import the SAS files
install.packages('haven')
install.packages('sas7bdat')
# Activate the `sas7bdat` library
library(sas7bdat)
library(haven)
# Read in the SAS data
#mySASData <- read.sas7bdat("example.sas7bdat")
#data <- read_sas("C:/temp/yoursasdataset.sas7bdat")

```
Once in a while, update all the packages by calling `update.packages()`.

## Read SAS `sas7bdat` data into R

### Use library `sas7bdat` in R

The [`sas7bdat`](https://www.rdocumentation.org/packages/sas7bdat/versions/0.5) library , developed by Matt Shotwell, works.

```{r eval=FALSE}
library(sas7bdat)
data <-read.sas7bdat("psu97ai.sas7bdat")
view(data)
``` 

[Two issues](https://stackoverflow.com/questions/30006822/read-sas-sas7bdat-data-into-r) with the method:         

  * It's slow
  * `sas7bdat` package is currently considered experimental.
  
### Use library `haven` package in R

The `haven::read_sas` method in `haven` package also works, and it is **faster** than the `read.sas7bdat` method (*[99.5% less time](https://stackoverflow.com/questions/30006822/read-sas-sas7bdat-data-into-r) than `sas7bdat::read.sas7bdat`)*. The difference is caused by saving a massive list of attributes for the files. 

```{r eval=FALSE}
library(haven)
data <- read_sas("C:/temp/yoursasdataset.sas7bdat")
View(data)
```
### Load Data files in R Studio 

You can manually load the data in the R Studio environment, choose `File >Import Dataset > From SAS...`, select the file location and click `import`. 

## RMarkdown Package

This website is created and powered by [RMarkdown](https://rmarkdown.rstudio.com/) developed by [Yihui Xie](https://yihui.name/en/), [Hugo](https://gohugo.io/), [Academic theme](https://sourcethemes.com/academic/#hero), and [Netlify](https://www.netlify.com/). 

The contents are created using [RMarkdown](https://rmarkdown.rstudio.com/) with guidance from the [Step-by-Step Instruction provided by ALISON PRESMANES HILL](https://alison.rbind.io/post/up-and-running-with-blogdown/),  ["Blogdown: Creating Websites with R Markdown"](https://bookdown.org/yihui/blogdown/) and [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

The website theme -- [Hugo-Academic](https://sourcethemes.com/academic/#hero)-- is a very good choice for the academics. It has a few advantages including: 

* Latex math rendering via simply enabling it in `config` toml file.
```{r eval=FALSE, indent="   "}
# In config\_default\param.toml
# Enable global LaTeX math rendering?
#   If false, you can enable it locally on a per page basis.
math = true
```
* rich widget system for customization. 

**Reference**

* Yihui Xie, J. J. Allaire, Garrett Grolemund (2018). [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Chapman & Hall/CRC. ISBN 9781138359338.  
* Yihui Xie, Amber Thomas, Alison Presmanes Hill (2017). [blogdown:  Creating Websites with R Markdown](https://bookdown.org/yihui/blogdown/). Chapman & Hall/CRC. ISBN 9780815363729.
* Yihui Xie (2016). [bookdown](https://bookdown.org/): [Authoring Books and Technical Documents with R Markdown](https://bookdown.org/yihui/bookdown/). Chapman and Hall/CRC. ISBN 9781138700109

### Blogdown & Hugo - Fix ToC (with Rmd)

The solution to fix Table of Contents (TOC) came from both [Create Websites with RMarkdown](https://bookdown.org/yihui/blogdown/default-template.html) and [Fubits@GIT](https://dadascience.design/post/r-some-tricks-when-working-with-blogdown-hugo-working-draft/) .
In the yaml header of your Rmd file, specific the following: 

```
-----
output: 
 blogdown::html_page:
  number_sections: true
  toc: true
  toc_depth: 3
  css: "/css/custom.css"
-----
```
Create and save the `custom.css` under the `\static\css` in the root directory. Using the css formating in [link](https://dadascience.design/post/r-some-tricks-when-working-with-blogdown-hugo-working-draft/), Your `custom.css` should look like: 

```css
/* Add some Heading / Title before the TOC */
#TOC:before{
    content: "Table of Content";
    font-family: 'Lato', sans-serif!important;
    font-weight:400;
    font-size: 26px;
}

/* Numbering suffix in TOC */
.toc-section-number:after{
    content: ". "
}
/* Numbering suffix in Body/Content */
.header-section-number:after{
    content: ". "
}
```
### Writing a Reproducible Paper in R Markdown

Paul C. Bauer created a template for [writing a reproducible scientific paper](https://drive.google.com/drive/folders/0BzJFjFv2bz2xVFFJLTVtQnp5V0U), the files can be downloaded [here](https://drive.google.com/drive/folders/0BzJFjFv2bz2xVFFJLTVtQnp5V0U). 
