---
title: R Notebook
author: Yuan Tian
date: '2019-08-16'
description: "Getting started with R, website creation and writing reproducible papers, tips and FAQs. "
slug: r-notebook
categories:
  - r
tags: [hugo,rmarkdown]
years: [2019]
image:
  caption: ''
  focal_point: ''
output: 
 blogdown::html_page:
  number_sections: true
  toc: true
  toc_depth: 3
  css: "/css/custom.css"
---

# **R Setup 101**

## R Packages for Basic R Installation or Update

For new installation, here is a list of most-used packages to install: 

```{r eval=FALSE}
install.packages('xlsx')
install.packages('rmarkdown')
install.packages("tinytex")
# install TinyTeX (must-do-this-step)
tinytex::install_tinytex()  
install.packages('tidyverse')
install.packages('ggplot2')
#table creation with specific formatting
install.packages('kableExtra')
#for writing reproducible research paper in R Markdown
install.packages(c("stargazer","plotly","knitr","bookdown"))
#To import the SAS files
install.packages('haven')
install.packages('sas7bdat')
install.packages('HSAUR3')
# Activate the `sas7bdat` library
library(sas7bdat)
library(haven)
# Read in the SAS data
#mySASData <- read.sas7bdat("example.sas7bdat")
#data <- read_sas("C:/temp/yoursasdataset.sas7bdat")

```
Once in a while, update all the packages by calling `update.packages()`.

## Read SAS `sas7bdat` data into R

### Use library `sas7bdat` in R

The [`sas7bdat`](https://www.rdocumentation.org/packages/sas7bdat/versions/0.5) library , developed by Matt Shotwell, works.

```{r eval=FALSE}
library(sas7bdat)
data <-read.sas7bdat("psu97ai.sas7bdat")
view(data)
``` 

[Two issues](https://stackoverflow.com/questions/30006822/read-sas-sas7bdat-data-into-r) with the method:         

  * It's slow
  * `sas7bdat` package is currently considered experimental.
  
### Use library `haven` package in R

The `haven::read_sas` method in `haven` package also works, and it is **faster** than the `read.sas7bdat` method (*[99.5% less time](https://stackoverflow.com/questions/30006822/read-sas-sas7bdat-data-into-r) than `sas7bdat::read.sas7bdat`)*. The difference is caused by saving a massive list of attributes for the files. 

```{r eval=FALSE}
library(haven)
data <- read_sas("C:/temp/yoursasdataset.sas7bdat")
View(data)
```
### Load Data files in R Studio 

You can manually load the data in the R Studio environment, choose `File >Import Dataset > From SAS...`, select the file location and click `import`. 

# **Create a Website using RMarkdown**

## RMarkdown, Hugo, Netlify

You can easily create beautiful static website powered by [RMarkdown](https://rmarkdown.rstudio.com/) developed by [Yihui Xie](https://yihui.name/en/), [Hugo](https://gohugo.io/), [Academic theme](https://sourcethemes.com/academic/#hero), and [Netlify](https://www.netlify.com/). 

The first step is to install blogdown and Hugo. Open RStudio , installing the Blogdown and Hugo dependencies:

```{r eval=FALSE,echo=TRUE}
 install.packages("blogdown")
 blogdown::install_hugo(version = "0.60.1", force = TRUE)
```

The contents are created using [RMarkdown](https://rmarkdown.rstudio.com/) with guidance from the [Step-by-Step Instruction provided by ALISON PRESMANES HILL](https://alison.rbind.io/post/up-and-running-with-blogdown/),  ["Blogdown: Creating Websites with R Markdown"](https://bookdown.org/yihui/blogdown/) and [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

Hugo has a variety of theme that you can choose from. For instance, [Hugo-Academic](https://sourcethemes.com/academic/#hero) is a very rich theme for the academics. A few advantages include: 

* Latex math rendering via simply enabling it in `config` toml file.
```{r eval=FALSE, indent="   "}
# In config\_default\param.toml
# Enable global LaTeX math rendering?
#   If false, you can enable it locally on a per page basis.
math = true
```
* rich widget system for customization. 

If you prefer a simple/minimal style, you can choose from the [minimalist](https://themes.gohugo.io/tags/minimalist/) themes. This website is created and powered by [Ezhil theme](https://themes.gohugo.io/ezhil/) and [hugo-xmin theme](https://themes.gohugo.io/hugo-xmin/), both themes were minimal/simple styles. 

Yihui Xie provided a step-by-step guide for the [hugo-xmin theme](https://themes.gohugo.io/hugo-xmin/) in the book [blogdown: Creating Websites with R Markdown](https://bookdown.org/yihui/blogdown/blogdown.pdf). It is very useful to finish the Hugo chapter to get a better understanding of its structure, rules of overriding, tips to add functionalities to your website (e.g., add math, add highlight.js, add custom.css for codes). 

Two most important folders related to website design are 

* *layouts*
  + partials: header.html, footer.html
  + _default: list.html,single.html,terms.html
* *static*
  + css:custom.css, fonts.css, styles.css

Modify the folders above in your own website folder but not the *exampleSite*, as they will override the same folder/files in the *theme* folder. This also will make it easy to maintain, migrate, and upgrade.


**Reference**

* Yihui Xie, J. J. Allaire, Garrett Grolemund (2018). [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Chapman & Hall/CRC. ISBN 9781138359338.  
* Yihui Xie, Amber Thomas, Alison Presmanes Hill (2017). [blogdown:  Creating Websites with R Markdown](https://bookdown.org/yihui/blogdown/). Chapman & Hall/CRC. ISBN 9780815363729.
* Yihui Xie (2016). [bookdown](https://bookdown.org/): [Authoring Books and Technical Documents with R Markdown](https://bookdown.org/yihui/bookdown/). Chapman and Hall/CRC. ISBN 9781138700109

## Blogdown & Hugo - Fix ToC (with Rmd)

The solution to fix Table of Contents (TOC) came from both [Create Websites with RMarkdown](https://bookdown.org/yihui/blogdown/default-template.html) and [Fubits@GIT](https://dadascience.design/post/r-some-tricks-when-working-with-blogdown-hugo-working-draft/) .
In the `yaml` header of your Rmd file, specific the following: 

```
-----
output: 
 blogdown::html_page:
  number_sections: true
  toc: true
  toc_depth: 3
  css: "/css/custom.css"
-----
```
Create and save the `custom.css` under the `\static\css` in the root directory. Using the css formating in [link](https://dadascience.design/post/r-some-tricks-when-working-with-blogdown-hugo-working-draft/), Your `custom.css` should look like: 

```css
/* Add some Heading / Title before the TOC */
#TOC:before{
    content: "Table of Content";
    font-family: 'Lato', sans-serif!important;
    font-weight:400;
    font-size: 26px;
}

/* Numbering suffix in TOC */
.toc-section-number:after{
    content: ". "
}
/* Numbering suffix in Body/Content */
.header-section-number:after{
    content: ". "
}
```
# **Write a Reproducible Paper using RMarkdown**

Unlike creating website using rmarkdown, writing a reproducible paper involves using another set of R packages, primarily those which advance markdown with *latex* features.

## Templates

### Paul's Template

Paul C. Bauer created a template for [writing a reproducible scientific paper](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3175518). All the files are downloadable from the [Shared Google Drive Folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). The [version 2](https://drive.google.com/drive/folders/1rzwN2jX_CW6OJFz5jrkObiUxGuMowR3U) has a simplified version without using the `header.tex`. 

I read the article, tested his codes and was able to reproduce the article. I ran into a few issues related to TinyTex installization, so just be mindful of the following steps:

* `install.packages("tinytex")` and `tinytex::install_tinytex()` to get the **latest** version of *TinyTex* (if bugs presents, re-install with the cmd `tinytex::reinstall_tinytex()`).
* When you *knit* your paper the first time, you should expect auto-installization of many *tex* packages and it could take some time depending on the number of *tex* packages you use in your document.
* also `install.packages("kableExtra")` for building [common complex tables with customizable styles](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf).  

### Reproducible APA manuscripts with RMarkdown

[Frederik Aust & Marius Barth](https://crsh.github.io/papaja_man/) wrote this comprehensive book covering lots of useful tricks and tips. This is a must-read if you seriously consider writing papers using Rmarkdown. 

## YAML options

There are a lot of advanced setting in the `YAML header`.  An example is provided below with some highlighted options:

* tex packages: `header-includes:`
* use `bookdown::pdf_document2: keep_tex: true` for pdf output creation.  
* `documentclass` can be the built-in `article` or your own customized `cls` file (e.g., `uofsthesis-cs.cls`).
* page setup: `geometry`, `fontsize`, and `linestretch`
* `bibliography`: put your `.bib` file in the same directory with your article.
* citation style: `csl`. You can search and download/create one(e.g., ACM or IEEE) from [zotero](https://www.zotero.org/styles?q=acm) or [Mendeley](https://csl.mendeley.com/searchByName/). Save it in the same directory with your article.
  - `nocite: |@Lander14` to include references currently not cited in the article. 

```
---
title: |
  | \vspace{1cm}\pagenumbering{gobble}This is my topic
  | A Literature Review \vspace{0.5cm}
author:  |
  | My Name
  | Department of Computer Science
date: |
  | `r gsub("^0", "", format(Sys.time(), "%d %B, %Y"))`
  |
abstract: this is my abstract
colorlinks: true
output: 
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
geometry: [top=0.85in,left=1in,right=1in, footskip=0.40in]
fontsize: [10pt,letterpaper]
linestretch: 1.1
documentclass: article
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{color}
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage[usenames,dvipsnames]{xcolor}
   - \usepackage[hang]{footmisc}
   - \renewcommand{\footnotesize}{\normalsize}
bibliography: references.bib
csl: transactions-on-management-information-systems.csl
suppress-bibliography: FALSE
always_allow_html: yes
link-citations: true
---
```
## Tricks - citations in a table

Cross-referencing tables or figures in a Rmarkdown article are straightforward and can be found in [Paul's article and template](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3175518).  However, citations within a table is very tricky in the `PDF`/`TEX` environment. It is easy to get it working in the HTML environment, but not with `pdf`, `pandoc`, and `latex`.

* Pandoc Markdown supports a citation extension to the basic markup, like `[@krycho]`. 

* For citation in a table in a `pdf` output file. [Rachael Lappan](https://github.com/rachaellappan) offered a [working solution](https://github.com/haozhu233/kableExtra/issues/214) for it.

>I had success with the bookdown text references - I'm not sure if you need to be making a bookdown book or if it will work with the bookdown package loaded.
>In my case, I defined the references first outside of a code block, e.g.
```
(ref:studies-table-ref1) Bogaert *et al.* (2011) [@BogaertVariabilityDiversityNasopharyngeal2011]
```
>then added them to the table
```
studies.table$Reference <- c("(ref:studies-table-ref1)", "(ref:studies-table-ref2)", # etc
```
>And kable interpreted them correctly.







# **FAQs and Tips**

## `save()` and `load()` datasets in R

For big datasets, importing the data and doing manipulaiton, aggregation and selection could take several minutes (or even longer), you might not want to start from scrach everytime you are working with the data. 

So, you hope you can save your dataset at a stage (e.g. pre-analyses). You can do that using the `save()` and `load()` functions in base-R. 

* [How to save and load datasets in R: an overview](https://www.r-bloggers.com/how-to-save-and-load-datasets-in-r-an-overview/)

## How to knit a document without rerunnning code

Dynamic Documents with and `knitr`: `knitr::options` give you multiple options which allow you to knit a document without re-running or displaying the code. 

Chunk options and package options can be found <https://yihui.org/knitr/options/>

A few highlights from the document: 

* `echo`: (TRUE; `logical` or `numeric`) whether to include R source code in the output file
* `eval`: (TRUE; `logical`) whether to evaluate the code chunk
* `collapse`: (FALSE; `logical`) collapse all the source and output blocks from one code chunk into a single block
* `include`: (TRUE; `logical`) whether to include the chunk output in the final output document; if include=FALSE, nothing will be written into the output document, **but the code is still evaluated** and plot files are generated. 

